Regression Cube Analysis of Cohort Study Data

### 2015-03-27 draft.tex
revised Hepatic Steatosis analysis section

Domain expert for the breast fat data set analysis is KH, a clinician (10 years of experience) with focus on epidemiological research.
%%
KH is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.

The hepatic steatosis data set is analyzed by UN, a data scientists responsible for prior analysis of the data with.
%%

TI, a statistician with focus on epidemiology (\com{TODO} years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
%%
\subsection{The Breast Fat Data Set}
The breast fat data set was compiled to find associations between parenchyma tissue proportion in the female breast towards other features in the data.
%%
The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
%%
Studies describe a four to five times increased risk of getting breast cancer for participants with a breast density above 50\% \cite{Mccormack2006}.

@@

In \cite{Niemann2014}, the authors show that women and men exhibit differences in the class distribution of the liver fat concentration.
%%
Also, for women they identify that there is a association between age and liver fat concentration.
%%
They compute an appropriate cut-off value of 52 years for which the class distribution is most homogeneous within the resulting subsets.
%%
Based on this observations, we perform our analysis on three sub-populations, $PartitionF$, $PartitionM$ and \emph{$F:age\>52$}.
%%
\subsection{Case 1: Hypothesis-free Analysis of the Breast Cancer Data Set}
%%
The analysis goal is to find relationships on the breast cancer data using features derived from the mammography analysis.
%%
Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
%%

\subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
%%
\com{1. Can we reconstruct relationships based on the regression cube?}

\com{2. How does the numeric target feature influence the result?}
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
\textbf{Extracted Hypotheses have to investigated further.}
\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
%%

>>>>

Domain expert for the breast fat data set analysis is KH, a clinician (10 years of experience) with focus on epidemiological research.
%%
\emph{KH} is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.

The hepatic steatosis data set is analyzed by \emph{UN}, a data scientists responsible for prior analysis of the data with.
%%

\emph{TI}, a statistician with focus on epidemiology (\com{TODO} years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
%%
\subsection{The Breast Fat Data Set}
The breast fat data set was compiled to find associations between parenchyma tissue proportion in the female breast towards other features in the data.
%%
The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
%%
Studies describe a four to five times increased risk of getting breast cancer for participants with a breast density above 50\% \cite{Mccormack2006}.

@@

In \cite{Niemann2014}, the authors show that women and men exhibit differences in the class distribution of the liver fat concentration.
%%
Also, for women they identify that there is a association between age and liver fat concentration.
%%
They compute an appropriate cut-off value of 52 years for which the class distribution is most homogeneous within the resulting subsets.
%%
Based on this observations, we perform our analysis on three sub-populations, \emph{females}, \emph{males} and \emph{females} older than 52 years.
%%
\subsection{Case 1: Hypothesis-free Analysis of the Breast Cancer Data Set}
%%
%%
The analysis goal is to find relationships on the breast cancer data using features derived from the mammography analysis.
%%
Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
%%

\subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
%%
We refer to each analysis step regarding to their belonging in the VA-Mantra (Subsec.~\ref{sec:Workflow}).
%%
The analysis goal is reproducing results based \cite{Niemann2014}.
%%
Therefore, \emph{UN} started the \textbf{AF} step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.
%%
The \emph{SI} step using the 3D cube representations hotspots at the end of the cube (Fig.~\ref{fig:application} a).
%%
\textbf{ZF} by slicing through the cube using the mouse input to inspect the hotspots using the mosaic plot (\textbf{DD}) revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
%%
Similar results present for analyzing the \emph{female} groups.
%%
\emph{UN} could reproduce most results, some features features exhibit lower correlations though, e.g., creatinine magnitudes.
%%
Relationships not described in \cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g. \emph{aspartate aminotransferase}.
%%
Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.

% %%
% The goal at the beginning of the analysis was to reproduce results based on the prior analysis.
% %%
% Subjects with a fatty liver could be identified using somatometric features, hepatic steatosis indicator-features as well as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
% %%
% The analysis was started with the target feature \emph{liver fat content} (\emph{mrt\_liverfat\_s2}), which was derived from the MRI data analysis.
% %%
% \emph{UN} therefore started the calculation using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subects.
% %%
% %%
% %%
% The data set containing all \emph{males} showed broad consensus with the expected results.
% %%
% The 3D-cube showed highlighted features at the end of the cube (Fig.~\ref{fig:application} a).
% %%
% Using the mouse-slicing allowed for fast navigation to the respective features.
% %%
% They include indicators hepatic steatosis ($R^2$:XX) as well as somatometric features, such as \emph{body mass index} (\emph{BMI}) and waist circumference.
% %%
% Feature correlations with \emph{gamma-glutamyl transpeptidase} (\emph{GGT}, indicating damaged livers if elevated) were found, but with low $R^2$ of XX.
%
% %%
% UN analyzed both \emph{female} groups in the same manner, concluding with the same findings regarding \emph{somatometric} features as well as \emph{hepatic steatosis} indicators.
% %%
% Using the mosaic plot, relationships incorporating enzyme used for the diagnosis of liver dysfunctions were found, e.g. aspartate aminotransferase.
% %%
%
% For most features, UN found at least tendencies using our approach.
% %%
% Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
%
% %Regression cube calculations were carried out for

\textbf{Analysis of non-discretized target feature.}
%%
Since our method can analyze numerical target features, as opposed to decision trees used by \emph{UN} before, the analysis was conducted again for the not-dichotomized target using the same formula.
%%
The 3D-cube representation already showed generally lower $R^2$ values, but since the analysis is now based on linear regression, the $R^2$ values can not be directly compared.
%%
Correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of XX for somatometric features as opposed to XX).
%%
We conclude that the bias introduced by dichotomizing the fat liver content enforces the findings towards liver diseases, while using the numerical features is less expressive.
%%

\textbf{Interleukin-6 correlation with liver fat.}
%%
During the analysis, one hotspot was always observed in the \emph{SI} and \emph{ZF} incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values (Fig.~\ref{fig:Application} b).
%%
The correlation was high for both the dichotomized and continuous target feature.
%%
In the literature, relations between liver cancer \cite{He2013} and also chronic liver diseases \cite{Streetz2003}.
%%
Strong effects of \emph{IL-6} with hepatic steatosis were also prior described with mice \cite{Hong2004}.
%%
The finding is subject of further analysis.
%%
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
\textbf{Extracted Hypotheses have to investigated further.}
\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
%%
